Lineage Analysis of CRISPR and shRNA Screens

### 2018-10-08 lineages_18Q3.Rmd

Load libraries.
```{r}
library(NMF)
library(ggpubr)
library(ggsignif)
library(ggfortify)
library(ggord)
library(rowr)
library(data.table)
library(CePa)
library(plyr)
library(tidyverse)
library(magrittr)
library(matrixStats)
library(parallel)
library(kableExtra)
library(gridExtra)
library(broom)
library(glmnet)
library(devtools)
library(reshape2)
library(caret)
library(bsselectR)
library(ComplexHeatmap)
library(circlize)
```

@@

crispr_ptmuts_lvlA_filt_plot <- ggplot(data = crispr_ptmuts_lvlA_filt, aes(x = Hugo_Symbol, y = Score)) +
  geom_hline(yintercept = 0, lty = 2, color = "darkgray") +
  geom_boxplot(mapping = aes(fill = Mutation_Status_Nonsilent), position = position_dodge(0.85)) +
  scale_fill_manual(values = crispr_ptmuts_lvlA_filt_color) +
  guides(color = FALSE) +
  theme(axis.text.x = element_text(angle = 45, hjust = 1), legend.direction = "horizontal", legend.justification = c(0.5, 0), legend.position = c(0.092, 0.8), legend.box.margin = margin(c(10, 10, 10, 10))) +
  labs(fill = "Mutation Status", y = "CERES Score", x = "Gene")
crispr_ptmuts_lvlA_filt_plot
# ggsave("./plots_18Q3/manuscript/crispr_ptmuts_lvlA_filt_plot.png", crispr_ptmuts_lvlA_filt_plot, device = "png", width = 20, height = 5, units = "in")

crispr_ptmuts_lvlA_filt_plot_faceted <- ggplot(data = crispr_ptmuts_lvlA_filt, aes(x = Mutation_Status_Nonsilent, y = Score)) +
  facet_wrap(~ Hugo_Symbol, drop = FALSE, ncol = length(unique(genes_lvlA_filt))) +
  geom_hline(yintercept = 0, lty = 2, color = "darkgray") +
  geom_boxplot(mapping = aes(fill = Mutation_Status_Nonsilent), position = position_dodge(0.85)) +
  scale_fill_manual(values = crispr_ptmuts_lvlA_filt_color) +
  guides(color = FALSE) +
  geom_text(data = label_text, mapping = aes(x = 1.5, y = 4.5, label = p.signif.adj)) +
  theme(legend.direction = "horizontal", legend.justification = c(0.5, 0), legend.position = c(0.25, 0.8), legend.box.margin = margin(c(10, 10, 10, 10)), axis.ticks.x = element_blank(), axis.text.x = element_blank(), axis.title.x = element_blank()) +
  labs(fill = "Mutation Status", y = "CERES Score", x = "Mutation Status")
crispr_ptmuts_lvlA_filt_plot_faceted
# ggsave("./plots_18Q3/manuscript/crispr_ptmuts_lvlA_filt_plot_faceted.png", crispr_ptmuts_lvlA_filt_plot_faceted, device = "png", width = 20, height = 4, units = "in")

crispr_ptmuts_lvlA_filt_plot_faceted_truncated <- ggplot(data = crispr_ptmuts_lvlA_filt, aes(x = Mutation_Status_Nonsilent, y = Score)) +
  facet_wrap(~ Hugo_Symbol, drop = FALSE, ncol = length(unique(genes_lvlA_filt))) +
  geom_hline(yintercept = 0, lty = 2, color = "darkgray") +
  coord_cartesian(ylim = c(-2.5, 2.5)) +
  geom_boxplot(aes(fill = Mutation_Status_Nonsilent), position = position_dodge(0.85)) +
  scale_fill_manual(values = crispr_ptmuts_lvlA_filt_color) +
  geom_text(data = label_text, mapping = aes(x = 1.5, y = 2.5, label = p.signif.adj)) +
  guides(color = FALSE) +
  theme(legend.direction = "horizontal", legend.justification = c(0.5, 0), legend.position = c(0.5, 0), legend.box.margin = margin(c(10, 10, 10, 10)), axis.ticks.x = element_blank(), axis.text.x = element_blank(), axis.title.x = element_blank()) +
  labs(fill = "Mutation Status", y = "CERES Score", x = "Mutation Status")
crispr_ptmuts_lvlA_filt_plot_faceted_truncated
# ggsave("./plots_18Q3/manuscript/crispr_ptmuts_lvlA_filt_plot_faceted_truncated.png", crispr_ptmuts_lvlA_filt_plot_faceted_truncated, device = "png", width = 20, height = 4, units = "in")
```

@@

```{r fig.height=5, fig.width=12}
g2p_indications <- filter(read.delim("./data_munging/data_mutation_associations_appended.csv", sep = "\t", header = TRUE), Evidence.Level == "A")
maf_g2p_indications <- filter(maf_raw, Genome_Change %in% g2p_indications$MutationName)
crispr_g2p_indications <- merge(maf_g2p_indications, crispr_data_ptmuts, by = c("Hugo_Symbol", "CCLE_Name", "Broad_ID"))
dup_g2p_indications <- filter(crispr_g2p_indications[, c("Hugo_Symbol", "CCLE_Name", "Broad_ID")] %>% group_by(Hugo_Symbol, CCLE_Name, Broad_ID) %>% tally(), n > 1)
crispr_g2p_indications <- merge(crispr_g2p_indications, dup_g2p_indications, by = c("Hugo_Symbol", "Broad_ID"), all.x = TRUE)
crispr_g2p_indications <- filter(crispr_g2p_indications, is.na(n))

@@

# ggsave("./plots_18Q3/manuscript/crispr_g2p_indications_plot_test.png", crispr_g2p_indications_plot, width = 12, height = 5, units = "in")
```

@@

```{r}
lvlA_signif_summ <- lvlA_signif %>% group_by(Hugo_Symbol, group_general_lineage_name, Mutation_Status_Nonsilent) %>% tally()

# write.table(lvlA_signif_summ, file = "~/Desktop/lvlA_signif_summ_20181005.tsv", quote = FALSE, sep = "\t")
```

>>>>

Load libraries.
```{r}
library(broom)
library(bsselectR)
library(caret)
library(CePa)
library(circlize)
library(ComplexHeatmap)
library(data.table)
library(devtools)
library(ggpubr)
library(ggsignif)
library(ggfortify)
library(ggord)
library(glmnet)
library(gridExtra)
# devtools::install_github("jeroen/jsonlite")
# install.packages("jsonlite")
library(jsonlite)
library(jsonlite)
library(kableExtra)
library(magrittr)
library(matrixStats)
library(NMF)
library(parallel)
library(plyr)
library(reshape2)
library(rowr)
library(tidyverse)
```

@@

crispr_ptmuts_lvlA_filt_plot <- ggplot(data = crispr_ptmuts_lvlA_filt, aes(x = Hugo_Symbol, y = Score)) +
  geom_hline(yintercept = 0, lty = 2, color = "darkgray") +
  geom_boxplot(mapping = aes(fill = Mutation_Status_Nonsilent), position = position_dodge(0.85)) +
  scale_fill_manual(values = crispr_ptmuts_lvlA_filt_color) +
  guides(color = FALSE) +
  theme(axis.text.x = element_text(angle = 45, hjust = 1), legend.direction = "horizontal", legend.justification = c(0.5, 0), legend.position = c(0.092, 0.8), legend.box.margin = margin(c(10, 10, 10, 10))) +
  labs(fill = "Mutation Status", y = "CERES Score", x = "Gene")
crispr_ptmuts_lvlA_filt_plot
# ggsave("./plots_18Q3/manuscript/crispr_ptmuts_lvlA_filt_plot.png", crispr_ptmuts_lvlA_filt_plot, device = "png", width = 20, height = 6, units = "in")

crispr_ptmuts_lvlA_filt_plot_faceted <- ggplot(data = crispr_ptmuts_lvlA_filt, aes(x = Mutation_Status_Nonsilent, y = Score)) +
  facet_wrap(~ Hugo_Symbol, drop = FALSE, ncol = length(unique(genes_lvlA_filt))) +
  geom_hline(yintercept = 0, lty = 2, color = "darkgray") +
  geom_boxplot(mapping = aes(fill = Mutation_Status_Nonsilent), position = position_dodge(0.85)) +
  coord_cartesian(ylim = c(-3, 5)) +
  scale_fill_manual(values = crispr_ptmuts_lvlA_filt_color) +
  guides(color = FALSE) +
  geom_text(data = label_text, mapping = aes(x = 1.5, y = 5, label = p.signif.adj)) +
  theme(legend.direction = "horizontal", legend.justification = c(0.5, 0), legend.position = c(0.5, 0), legend.box.margin = margin(c(10, 10, 10, 10)), axis.ticks.x = element_blank(), axis.text.x = element_blank(), axis.title.x = element_blank()) +
  labs(fill = "Mutation Status", y = "CERES Score", x = "Mutation Status")
crispr_ptmuts_lvlA_filt_plot_faceted
# ggsave("./plots_18Q3/manuscript/crispr_ptmuts_lvlA_filt_plot_faceted.png", crispr_ptmuts_lvlA_filt_plot_faceted, device = "png", width = 20, height = 5, units = "in")

crispr_ptmuts_lvlA_filt_plot_faceted_truncated <- ggplot(data = crispr_ptmuts_lvlA_filt, aes(x = Mutation_Status_Nonsilent, y = Score)) +
  facet_wrap(~ Hugo_Symbol, drop = FALSE, ncol = length(unique(genes_lvlA_filt))) +
  geom_hline(yintercept = 0, lty = 2, color = "darkgray") +
  coord_cartesian(ylim = c(-2.5, 2.5)) +
  geom_boxplot(aes(fill = Mutation_Status_Nonsilent), position = position_dodge(0.85)) +
  scale_fill_manual(values = crispr_ptmuts_lvlA_filt_color) +
  geom_text(data = label_text, mapping = aes(x = 1.5, y = 2.5, label = p.signif.adj)) +
  guides(color = FALSE) +
  theme(legend.direction = "horizontal", legend.justification = c(0.5, 0), legend.position = c(0.5, 0), legend.box.margin = margin(c(10, 10, 10, 10)), axis.ticks.x = element_blank(), axis.text.x = element_blank(), axis.title.x = element_blank()) +
  labs(fill = "Mutation Status", y = "CERES Score", x = "Mutation Status")
crispr_ptmuts_lvlA_filt_plot_faceted_truncated
# ggsave("./plots_18Q3/manuscript/crispr_ptmuts_lvlA_filt_plot_faceted_truncated.png", crispr_ptmuts_lvlA_filt_plot_faceted_truncated, device = "png", width = 20, height = 5, units = "in")
```

@@

```{r fig.height=5, fig.width=20}
g2p_indications <- filter(read.delim("./data_munging/data_mutation_associations_appended.csv", sep = "\t", header = TRUE), Evidence.Level == "A")
maf_g2p_indications <- filter(maf_raw, Genome_Change %in% g2p_indications$MutationName)
crispr_g2p_indications <- merge(maf_g2p_indications, crispr_data_ptmuts, by = c("Hugo_Symbol", "CCLE_Name", "Broad_ID"))
dup_g2p_indications <- filter(crispr_g2p_indications[, c("Hugo_Symbol", "CCLE_Name", "Broad_ID")] %>% group_by(Hugo_Symbol, CCLE_Name, Broad_ID) %>% tally(), n > 1)
crispr_g2p_indications <- merge(crispr_g2p_indications, dup_g2p_indications, by = c("Hugo_Symbol", "Broad_ID"), all.x = TRUE)
crispr_g2p_indications <- filter(crispr_g2p_indications, is.na(n))

@@

crispr_g2p_indications_muts_plot <- ggplot(crispr_g2p_indications, aes(x = Protein_Change, y = Score)) +
  facet_grid(~ Hugo_Symbol, scales = "free_x", space = "free") +
  geom_hline(yintercept = 0, lty = 2, color = "darkgray") +
  geom_boxplot(alpha = 0.5, color = "lightgray", outlier.shape = NA) +
  geom_jitter(width = 0.1, mapping = aes(color = group_general_lineage_name), size = 1) +
  theme_light() +
  theme(legend.title = element_blank(), axis.text.x = element_text(angle = 90, hjust = 1)) +
  labs(x = "Gene", y = "CERES Score")
crispr_g2p_indications_muts_plot

# ggsave("./plots_18Q3/manuscript/crispr_g2p_indications_plot_test.png", crispr_g2p_indications_plot, width = 12, height = 5, units = "in")
# ggsave("./plots_18Q3/manuscript/crispr_g2p_indications_muts_plot_test.png", crispr_g2p_indications_muts_plot, width = 20, height = 5, units = "in")
```

```{r}
g2pA <- read.delim("./data_munging/g2p_full_level_A_dataset_v2_20181008.tsv", sep = "\t", header = TRUE, na.strings = "-1", quote = "")
# g2pA[g2pA == "-1"] <- NA

g2pA$JSON[1]
fromJSON(as.character(g2pA$JSON[[1]]))
```

```{r}
test <- '{"feature_names": "ALK positive", "tags": "test", "genes": "ALK"}'

fromJSON(test)
```

```
{
'feature_names': 'ALK positive',
'jax': '{\"responseType\": \"sensitive\", 
        \"efficacyEvidence\": \"Zykadia (ceritinib) is FDA approved for the treatment of ALK-positive metastatic non-small cell lung cancer (NSCLC) in patients previously treated with crizotinib (FDA.gov).\",
        \"molecularProfile\": {\"profileName\": \"ALK positive\",
                               \"id\": 1772},
        \"evidenceType\": \"Actionable\",
        \"indication\": {\"source\": \"DOID\",
                         \"id\": 3908,
                         \"name\": \"non-small cell lung carcinoma\"},
        \"references\": [{\"url\": \"http://www.fda.gov/\",
                          \"id\": 286,
                          \"pubMedId\": null,
                          \"title\": \"FDA.gov\"}],
        \"approvalStatus\": \"FDA approved\",
        \"therapy\": {\"id\": 789,
                      \"therapyName\": \"Ceritinib\"},
        \"id\": 1481}',
'tags': [],
'genes': ['ALK'],
'source_url': 'https://ckb.jax.org/therapy/show/789',
'source': 'jax',
'dev_tags': [],
'gene_identifiers': [{'symbol': 'ALK',
                      'entrez_id': '238',
                      'ensembl_gene_id': 'ENSG00000171094'}],
'association': {'drug_labels': 'CERITINIB',
                'description': 'Zykadia (ceritinib) is FDA approved for the treatment of ALK-positive metastatic non-small cell lung cancer (NSCLC) in patients previously treated with crizotinib (FDA.gov).',
                'publication_url': 'http://www.fda.gov/',
                'source_link': 'https://ckb.jax.org/molecularProfile/show/1772',
                'variant_name': [''],
                'phenotypes': [{'source': 'http://purl.obolibrary.org/obo/DOID_3908',
                                'term': 'non-small cell lung carcinoma',
                                'description': 'non-small cell lung carcinoma',
                                'family': 'lung cancer',
                                'id': 'DOID:3908'}],
                'evidence': [{'info': {'publications': ['http://www.fda.gov/']},
                              'evidenceType': {'sourceName': 'jax'},
                              'description': 'sensitive'}],
                              'environmentalContexts': [{'term': 'CERITINIB',
                                                          'description': 'CERITINIB',
                                                          'taxonomy': {'kingdom': 'Chemical entities',
                                                                       'direct-parent': 'Phenylpiperidines',
                                                                       'class': 'Organoheterocyclic compounds',
                                                                       'subclass': 'Piperidines',
                                                                       'superclass': 'Organic compounds'},
                                                          'source': 'http://rdf.ncbi.nlm.nih.gov/pubchem/compound',
                                                          'usan_stem': 'tyrosine kinase inhibitors',
                                                          'toxicity': 'There is not currently any data on carcinogenicity, effect on human fertility, or on early embryonic development. However, based on its mechanism of action, ceritinib may cause fetal harm when administered to pregnant women and should therefore be administered with effective contraception during treatment. Diarrhea, nausea, vomiting, or abdominal pain occurred in 96% of 255 patients including severe cases in 14% of patients. Drug-induced hepatotoxicity also occurred in 27% of 255 patients, presenting as alanine aminotransferase (ALT) levels greater than 5 times the upper limit of normal (ULN). Severe, life-threatening, or fatal interstitial lung disease (ILD)/pneumonitis, hyperglycaemia, and bradycardia have also been reported.',
                                                          'approved_countries': ['Canada',
                                                                                 'US'],
                                                          'id': 'CID57379345'}],
                'evidence_label': 'A',
                'response_type': 'sensitive',
                'evidence_level': 1},
'features': [{'provenance_rule': 'default_feature',
              'end': 29484957,
              'description': 'ALK  positive',
              'links': ['http://reg.genome.network/refseq/RS000050',
                        'http://reg.genome.network/refseq/RS000026',
                        'http://reg.genome.network/allele/CA425438613',
                        'http://reg.genome.network/refseq/RS000002',
                        'http://myvariant.info/v1/variant/chr2:g.29484957C>G?assembly=hg19',
                        'http://reg.genome.network/refseq/RS001597'],
              'sequence_ontology': {'hierarchy': ['SO:0000400',
                                                  'SO:0000733',
                                                  'SO:0000401',
                                                  'SO:0000119',
                                                  'SO:0000123',
                                                  'SO:0000125'],
                                    'soid': 'SO:0000475',
                                    'root_name': 'sequence_attribute',
                                    'name': 'positively_autoregulated',
                                    'root_soid': 'SO:0000400'},
              'provenance': ['http://myvariant.info/v1/query?q=ALK  positive',
                             'http://reg.genome.network/allele?hgvs=NC_000002.11%3Ag.29484957C%3EG'], 
              'start': 29484957,
              'synonyms': ['NC_000002.11:g.29484957C>G',
                           'NC_000002.10:g.29338461C>G',
                           'LRG_488:g.664476G>C',
                           'chr2:g.29484957C>G',
                           'NG_009445.1:g.664476G>C',
                           'NC_000002.12:g.29262091C>G',
                           'CM000664.1:g.29484957C>G',
                           'CM000664.2:g.29262091C>G'],
              'biomarker_type': 'positive',
              'referenceName': 'GRCh37',
              'geneSymbol': 'ALK',
              'alt': 'G',
              'ref': 'C',
              'chromosome': '2',
              'name': 'ALK  positive'}]
}
```

@@

```{r}
lvlA_signif_summ <- lvlA_signif %>% group_by(Hugo_Symbol, group_general_lineage_name, Mutation_Status_Nonsilent) %>% tally()
lvlA_signif_summ
# write.table(lvlA_signif_summ, file = "~/Desktop/lvlA_signif_summ_20181005.tsv", quote = FALSE, sep = "\t")
```
